---
title: "Library Survey Data"
author: "Janet Bang"
date: "January 20, 2025"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    extra_dependencies: ["float"]
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
# echo = F hides code # message = F hides messages
# fig.pos and out.extra = there to avoid the figures from floating around on the page when printing to pdf
knitr::opts_chunk$set(echo = F, message = F, fig.pos = "!H", out.extra = "") 
```


# Prep
## Load libraries and set theme
```{r}
library(tidyverse)
library(psych)
library(readxl)

theme_set(theme_minimal())
```


## Read in data
```{r}
# variables excluded to ensure anonymity of participants: age, library branch, years of experience
df <- read.csv("libsurvey_df_n70_share.csv") %>% 
  mutate(Q10_cc_understanding_11_TEXT = na_if(Q10_cc_understanding_11_TEXT, ""),
         Q23_language_request_11_TEXT = na_if(Q23_language_request_11_TEXT, ""),
         Q24_topic_request_8_TEXT2 = na_if(Q24_topic_request_8_TEXT2, ""),
         Q26_share_nonEngbook_7_TEXT = na_if(Q26_share_nonEngbook_7_TEXT, ""),
         Q29_share_Engbooks_7_TEXT = na_if(Q29_share_Engbooks_7_TEXT, ""),
         Q20_often_request = factor(Q20_often_request,
                                    levels = c("Never or rarely", "Sometimes", "Often", "Most of the time", "I'm not sure")),
         Q21_diff_library = factor(Q21_diff_library,
                                    levels = c("Never or rarely", "Sometimes", "Often", "Most of the time", "I'm not sure")))
```



# Q10 - Languages library staff are currently comfortable understanding
```{r}
# clean up column
df_q10_lang_und <- df %>% 
  dplyr::select(Keep, id, Q10_cc_understanding, Q10_cc_understanding_11_TEXT) %>%
  separate(Q10_cc_understanding, c("lang1", "lang2", "lang3", "lang4", "lang5", 
                                "lang6", "lang7", "lang8", "lang9", "lang10"), 
           sep = ",") %>% 
  mutate(Q10_cc_understanding_11_TEXT = str_replace(Q10_cc_understanding_11_TEXT, "a little bit of spanish and french", "French")) %>% 
  mutate_at(vars(lang1, lang2, lang3, lang4, lang5, lang6, lang7, 
                  lang8, lang9, lang10), na_if, "Other (please specify)") %>% 
  mutate_at(vars(lang1, lang2, lang3, lang4, lang5, lang6, lang7, 
                  lang8, lang9, lang10), na_if, "") %>% 
  pivot_longer(cols = lang1:Q10_cc_understanding_11_TEXT, 
               values_to = "languages") %>%
  na.omit()


# create counter for languages
df_q10_seq <- df_q10_lang_und %>% 
  group_by(Keep, id, name) %>% 
  distinct(Keep, name) %>% 
  group_by(Keep, id) %>% 
  count() %>% 
  rename("lang_num" = "n")


# combine data
df_q10_lang_und2 <- df_q10_lang_und %>% 
  full_join(df_q10_seq, by = c("id", "Keep")) %>% 
  mutate(skill = "understanding", 
         dev_time = "current")

# count by language
lang_und <- df_q10_lang_und2 %>% 
  group_by(Keep, languages) %>% 
  count(languages, sort = T) %>% 
  mutate(prop = round((n/70)*100, 0))

lang_und

# count of languages by individual
df_q10_seq %>% 
  group_by(Keep, lang_num) %>% 
  count(lang_num, sort = T) %>% 
  mutate(prop = round((n/70)*100, 0))
```


# Table 2 - Responses regarding borrowing patterns for non-English books
## Q19 - Families know that the library offers non-English books
```{r}
df %>% 
  group_by(Q19_aware_nonEnglish) %>% 
  count() %>% 
  mutate(prop = n/69) # 1 person did not respond
```


## Q20 - How often do you feel families ask for non-English books
```{r}
df %>% 
  group_by(Q20_often_request) %>% 
  count() %>% 
  mutate(prop = n/69) # 1 person did not respond
```


## Q21 - Request for non-English book from different library
```{r}
df %>% 
  group_by(Q21_diff_library) %>% 
  count() %>% 
  mutate(prop = n/69) # 1 person did not respond
```




# Figure 1
## Q23 - Family requests for non-English books - LANGUAGE
```{r}
# clean up column
df_q23_lang_requests <- df %>% 
  dplyr::select(id, Q23_language_request, Q23_language_request_11_TEXT) %>%
  separate(Q23_language_request, c("lang1", "lang2", "lang3", "lang4", "lang5", 
                                "lang6", "lang7", "lang8", "lang9", "lang10", 
                                "lang11", "lang12", "lang13", "lang14", "lang15", 
                                "lang16", "lang17", "lang18"), 
           sep = ",") %>% 
  mutate_at(vars(lang1, lang2, lang3, lang4, lang5, lang6, lang7, 
                  lang8, lang9, lang10, lang11, lang12, lang13,
                  lang14, lang15, lang16, lang17, lang18), na_if, "Other (please specify)") %>% 
  mutate_at(vars(lang1, lang2, lang3, lang4, lang5, lang6, lang7, 
                  lang8, lang9, lang10, lang11, lang12, lang13,
                  lang14, lang15, lang16, lang17, lang18), na_if, "none") %>% 
  mutate_at(vars(lang1, lang2, lang3, lang4, lang5, lang6, lang7, 
                  lang8, lang9, lang10, lang11, lang12, lang13,
                  lang14, lang15, lang16, lang17, lang18), na_if, "") %>% 
  pivot_longer(cols = lang1:Q23_language_request_11_TEXT, 
               values_to = "languages") %>% 
  na.omit() %>% 
  filter(languages != "Braille" & languages != "English" & 
         languages != "none")


# create counter for languages
df_q23_seq <- df_q23_lang_requests %>% 
  group_by(id, name) %>% 
  distinct(name) %>% 
  group_by(id) %>% 
  count() %>% 
  rename("lang_num" = "n")


# combine data
df_q23_lang_requests2 <- df_q23_lang_requests %>% 
  full_join(df_q23_seq, by = c("id"))


# count by language = 23 non-English languages
count_by_lang <- df_q23_lang_requests2 %>% 
  group_by(languages) %>% 
  count(languages, sort = T) %>% 
  ungroup() %>% 
  mutate(total_responses = sum(n),
         prop_outof_responses = round((n /total_responses), 3), 
         languages = factor(languages))


# count number of languages
df_q23_lang_requests2 %>% 
  distinct(languages) %>% 
  nrow()


# count number of responses - confirming number of responses n = 68
df_q23_lang_requests2 %>% 
  distinct(id) %>% 
  nrow()


# plot - to visualize
# ggplot(df_q23_lang_requests2, aes(x = fct_infreq(languages))) + 
#   geom_bar(aes(fill = fct_infreq(languages))) + 
#   theme(legend.position = "none", 
#         text = element_text(size = 35),
#         axis.text.x = element_text(angle = 45, hjust=1)) + 
#   labs(title = "Number of Library Staff Who Heard Requests", x = "languages", y = "")

# ggplot(count_by_lang, aes(x = prop_outof_responses, fill = languages)) + 
#   geom_bar() + 
#   labs(title = "Proportion of Responses per Language", x = "languages", y = "")
```


## Align with census data
```{r}
df_pop <- read.csv("santaclara_libsurvey_forR_share.csv") %>% 
  mutate(language = "language", 
         prop_outof_responses = -prop_outof_responses) %>% 
  pivot_longer(cols = c(prop_outof_responses, prop_pop, prop_nonEng),
               values_to = "proportion") %>%
  filter(name != "prop_pop") %>% 
  mutate(language_clean = fct_reorder(language_clean, desc(our_ranking)))


ggplot(df_pop, aes(x = proportion, y = language_clean, fill = name)) + 
  geom_col() + 
  theme(legend.position = "none",
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(title = "", x = "Proportion", y = "") + 
  xlim(-0.4, 0.4) +
  annotate("text", x = -.25, y = 16, size = 9, label = "Proportion out of \n Non-English Languages \nRequested") +
  annotate("text", x = .25, y = 16, size = 9, label = "Proportion out of \nNon-English Speakers \nin County")


ggsave("prop_lang.pdf", height = 10, width = 14, dpi = 300) # negative sign for -0.4 and -0.2 removed using illustrator


# inferential stats
# prop_outof_responses = prop responses out of all non-English responses
# percent_nonEng = prop out of non-English speakers
df_pop_wide <- read.csv("santaclara_libsurvey_forR_share.csv") %>% 
  select(language_clean, prop_outof_responses, prop_nonEng)

# ggplot(df_pop_wide, aes(prop_outof_responses, prop_nonEng, color = language_clean)) + 
#   geom_point(size = 3)

cor.test(df_pop_wide$prop_outof_responses, df_pop_wide$prop_nonEng, method = 'spearman', exact = F)

```


# Figure 2 - Q24 - Family requests for non-English books - TOPICS
```{r}
# write-in responses include: 
# 1) Translations of English-language classics
# 2) Stories/Fairytales
# 3) Not sure/Have not heard any requests
# 4) Language is requested more than topic
# 5) Any topic

# import in closed-ended question and write-in responses
df_q24_topic_requests <- df %>% 
  dplyr::select(id, Q24_topic_request, Q24_topic_request_8_TEXT2) %>%
  mutate(Q24_topic_request = str_replace_all(Q24_topic_request, "e.g.,", "for ex."), 
         Q24_topic_request = str_replace_all(Q24_topic_request, "bedtime,", "bedtime/"), 
         Q24_topic_request = str_replace_all(Q24_topic_request, "sibling,", "sibling/")) %>% 
  separate(Q24_topic_request, c("topic1", "topic2", "topic3", "topic4", "topic5", 
                              "topic6", "topic7", "topic8", "topic9", "topic10",
                              "topic11", "topic12", "topic13", "topic14", "topic15"), # separating commas for each selected topic
           sep = ",") %>% 
  mutate_at(vars(topic1, topic2, topic3, topic4, topic5, topic6, topic7, 
                  topic8, topic9, topic10, topic11, topic12, topic13,
                  topic14, topic15), na_if, "") %>% 
  pivot_longer(cols = topic1:Q24_topic_request_8_TEXT2, 
               values_to = "topics") %>% 
  na.omit() %>% 
  filter(topics != "Other (please specify)") %>% 
  dplyr::select(-name)
  


# count number of library staff with responses
df_q24_topic_requests %>% 
  filter(topics != "" & topics != "Any topic" & topics != "Language is requested more than topic" & 
         topics != "Not sure/Have not heard any requests") %>% 
  distinct(id) %>% 
  nrow()


# count by topic - number of responses
df_topics_plot <- df_q24_topic_requests %>% 
  group_by(topics) %>% 
  filter(topics != "" & topics != "Any topic" & topics != "Language is requested more than topic" & 
         topics != "Not sure/Have not heard any requests") %>% 
  count(topics, sort = T) %>% 
  ungroup() %>% 
  mutate(total_responses = sum(n),
    prop_outof_responses = round(n/total_responses, 3), 
         topics = factor(topics))

df_topics_plot


# plot
ggplot(df_topics_plot, aes(x = reorder(topics, prop_outof_responses), y = prop_outof_responses, fill = topics)) + 
  geom_col() + 
  geom_text(aes(label = prop_outof_responses), size = 8, hjust = "left") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1), 
        text = element_text(size = 30)) +
  labs(x = "Topics", y = "Proportion out of Responses") + 
  coord_flip() + 
  scale_fill_manual(values = c("coral2", "mediumpurple1", "darkgrey", "darkgrey", "darkgrey", 
                               "darkgrey", "darkgrey", "darkgrey", "darkgrey", "springgreen2", 
                               "dodgerblue", "darkgrey", "darkgrey", "darkgrey", "darkgrey", 
                               "goldenrod2"))

ggsave("prop_topics.pdf", height = 10, width = 22, dpi = 300)

```



# Table 3
## Q26 - Family reasons to share NON-ENGLISH books
```{r}
df_q26_reasons_noneng <- df %>% 
  dplyr::select(id, Q26_share_nonEngbook, Q26_share_nonEngbook_7_TEXT) %>%
  separate(Q26_share_nonEngbook, c("reason1", "reason2", "reason3", "reason4", "reason5", 
                              "reason6", "reason7"), 
           sep = ",") %>% 
  mutate_at(vars(reason1, reason2, reason3, reason4, reason5, reason6, reason7), na_if, "") %>% 
  pivot_longer(cols = reason1:Q26_share_nonEngbook_7_TEXT, 
               values_to = "reasons") %>% 
  na.omit() %>% 
  filter(reasons != "none" & reasons != "Other (please elaborate)")


# Reasons - only 6 categories - added these write-in reasons by hand
df_q26_reasons_noneng_6 <- df_q26_reasons_noneng %>% 
  filter(name != "Q26_share_nonEngbook_7_TEXT")

# Other reasons - only 1 new category created based on these responses because the others could be grouped into existing categories by same participant
# for paper - created new category for "Build family connections" (n = 2)
df_q26_reasons_noneng_other <- df_q26_reasons_noneng %>% 
  filter(name == "Q26_share_nonEngbook_7_TEXT")


df_q26_reasons_noneng_count <- df_q26_reasons_noneng %>% 
  group_by(reasons) %>% 
  count(reasons, sort = T) %>% 
  mutate(prop = n/67) # 1 person did not respond, 1 person did not give a response that could be coded "I think you have it backward...", and 1 person wrote 'none'

df_q26_reasons_noneng_count



# plot for visualization
# ggplot(df_q26_reasons_noneng_6, aes(x = reasons, fill = reasons)) + 
#   geom_bar() + 
#   theme(legend.position = "none", 
#         axis.text.x = element_text(angle = 90)) + 
#   labs(title = "REASONS for seeking NON-ENGLISH language")
```



## Q29 - Family reasons to share ENGLISH LANGUAGE books
```{r}
df_q29_reasons_eng <- df %>% 
  dplyr::select(id, Q29_share_Engbooks, Q29_share_Engbooks_7_TEXT) %>%
  mutate(Q29_share_Engbooks = str_replace_all(Q29_share_Engbooks, "matter,", "matter")) %>% 
  separate(Q29_share_Engbooks, c("reason1", "reason2", "reason3", "reason4", "reason5", 
                              "reason6", "reason7"), 
           sep = ",") %>% 
  mutate_at(vars(reason1, reason2, reason3, reason4, reason5, reason6, reason7), na_if, "") %>% 
  pivot_longer(cols = reason1:Q29_share_Engbooks_7_TEXT, 
               values_to = "reasons") %>% 
  na.omit() %>% 
  filter(reasons != "none" & reasons != "blank" & reasons != "Other (please elaborate)")


# Reasons - only 6 categories - added these write-in reasons by hand
df_q29_reasons_eng_6 <- df_q29_reasons_eng %>% 
  filter(name != "Q29_share_Engbooks_7_TEXT")


# Other reasons - only 1 new category created based on these responses because the others could be grouped into existing categories by same participant

# created one for "Be successful in school"
df_q29_reasons_eng_other <- df_q29_reasons_eng %>% 
  filter(name == "Q29_share_Engbooks_7_TEXT")


# count by topic
df_q29_reasons_eng %>% 
  group_by(reasons) %>% 
  count(reasons, sort = T) %>% 
  mutate(prop = n/68) # 1 person did not respond and was blank, 1 person said "none"


# plot for visualization
# ggplot(df_q29_reasons_eng_6, aes(x = reasons, fill = reasons)) + 
#   geom_bar() + 
#   theme(legend.position = "none", 
#         axis.text.x = element_text(angle = 90)) + 
#   labs(title = "REASONS for seeking ENGLISH language")
```



# Q30 - 5 things to recommend for families with young children
## Read in data
```{r}
# the column of "order" groups responses by participant, but differs from participant id above
df_q30 <- read_excel("libsurvey_q30_data_share.xlsx") %>% 
  dplyr::select(order, drop, Q30_clean_1, Q30_clean_2, Q30_clean_3, Q30_clean_4, Q30_clean_5) %>% 
  rename("First" = Q30_clean_1, "Second" = Q30_clean_2, "Third" = Q30_clean_3, "Fourth" = Q30_clean_4, "Fifth" = Q30_clean_5) %>% 
  mutate(order = factor(order),
         First = factor(First), 
         Second = factor(Second), 
         Third = factor(Third), 
         Fourth = factor(Fourth), 
         Fifth = factor(Fifth)) %>% 
  pivot_longer(cols = c(First, Second, Third, Fourth, Fifth), names_to = "thing_num", values_to = "codes") %>% 
  mutate(thing_num = factor(thing_num, levels = c("First", "Second", "Third", "Fourth", "Fifth"))) %>% 
  filter(drop != "drop") %>% # n = 6 who did not provide any responses
  dplyr::select(-drop) %>% 
  mutate(codes2 = ifelse(codes == "diversity of characters or authors", "content",
                 ifelse(codes == "book about specific cultures", "content", 
                 ifelse(codes == "classics/popularity", "content",
                 ifelse(codes == "content", "content",  
                 ifelse(codes == "age", "age", 
                 ifelse(codes == "availability", "availability", 
                 ifelse(codes == "book format", "book format", 
                 ifelse(codes == "child or family interests", "child or family interests", 
                 ifelse(codes == "condition", "condition", 
                 ifelse(codes == "illustrations", "illustrations", 
                 ifelse(codes == "language", "language", 
                 ifelse(codes == "length of book/reading level", "length of book/reading level", 
                 ifelse(codes == "staff familiarity", "staff familiarity", 
                 ifelse(codes == "gender", "blank_or_exclude", "blank_or_exclude")))))))))))))), # initial code of gender excluded because response deemed too vague
         codes2 = factor(codes2, levels = c("blank_or_exclude", "condition", "availability",
                                            "staff familiarity", "illustrations", "language", "book format", "age", 
                                            "length of book/reading level", "child or family interests", "content"))) 
# codes 2 above is ordered least to most frequent across whole sample

# number of participants; n = 64
# Note: n = 6 do not have codable responses (blank or exclude) for all 5 factors, so data reflects 64 participants
df_q30 %>% 
  group_by(order) %>% 
  distinct(order) %>% 
  nrow()

# # examine code per participant
# ggplot(df_q30, aes(order, fill = codes2)) + 
#   geom_bar()
```


## Calculate total per category across all five responses - Eventually removes duplicates within participant
```{r}
# remove blank or references to language
df_q30_excluded <- df_q30 %>% 
  filter(codes2 != "blank_or_exclude" & codes2 != "language") %>% 
  group_by(order) %>% 
  count()

psych::describe(df_q30_excluded$n, fast = T)


# remove duplicates
df_q30_distinct <- df_q30 %>% 
  select(-c(codes)) %>% 
  filter(codes2 != "blank_or_exclude" & codes2 != "language") %>% 
  distinct(order, codes2, .keep_all = T)

# count per participant
df_q30_distinct_count_id <- df_q30_distinct %>% 
  group_by(order) %>% 
  count()

psych::describe(df_q30_distinct_count_id$n, fast = T)


# count per code
df_q30_distinct_count_code <- df_q30_distinct %>% 
  group_by(codes2) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  ungroup() %>% 
  mutate(total = sum(n), 
         prop = round(n / total, 2))

df_q30_distinct_count_code


# visualize
# ggplot(df_q30_distinct_count_code, aes(codes2, prop, fill = codes2)) + 
#   geom_col(aes(x = reorder(codes2, -prop), y = prop)) +
#   theme(legend.position = "none", 
#         # text = element_text(size = 35),
#         axis.text.x = element_text(angle = 45, hjust=1))

```



## Visualize position for each code
```{r}
# number of responses per thing_num
num_resp_per_thing <- df_q30_distinct %>% 
  group_by(thing_num) %>% 
  count() %>% 
  rename(n_per_order = n)


# create df
df_q30_distinct_with_order <- df_q30 %>% 
  select(-codes) %>% 
  filter(codes2 != "blank_or_exclude" & codes2 != "language") %>% 
  distinct(order, codes2, .keep_all = T) %>% # .keep_all retains thing_num
  group_by(thing_num, codes2) %>% 
  count() %>% 
  rename(n_code = n) %>% 
  full_join(num_resp_per_thing, by = "thing_num") %>% 
  mutate(prop_code_per_order = round((n_code / n_per_order), 3)) %>% 
  mutate(codes2 = factor(codes2, levels = c("content", "child or family interests", "length of book/reading level", 
                                            "age", "book format", "illustrations", "staff familiarity",
                                            "availability", "condition")))
  
  
# bar graphs - stacked with percentage labels
# ggplot(df_q30_distinct_with_order, aes(thing_num, prop_code_per_order, fill = codes2)) + 
#   geom_col() + 
#   theme(text = element_text(size = 35),
#         axis.text.x = element_text(angle = 45, hjust=1)) + 
#   geom_text(aes(label = prop_code_per_order),
#             position=position_stack(vjust = 0.5))


# bar graphs - with code labels
ggplot(df_q30_distinct_with_order, aes(thing_num, prop_code_per_order, fill = codes2)) +
  geom_col(color = "black") + 
  theme(text = element_text(size = 30),
        axis.text.x = element_text(angle = 30, hjust=1), 
        legend.position = "none") + 
  geom_text(aes(label = codes2),
            size = 6,
            position=position_stack(vjust = 0.5)) +
  labs(x = "Order", y = "Proportion of Participants") +
  annotate("text", x = 1, y = 1.1, label = "n = 50", size = 7) + 
  annotate("text", x = 2, y = 1.1, label = "n = 46", size = 7) +  
  annotate("text", x = 3, y = 1.1, label = "n = 34", size = 7) + 
  annotate("text", x = 4, y = 1.1, label = "n = 18", size = 7) + 
  annotate("text", x = 5, y = 1.1, label = "n = 5", size = 7) + 
  scale_fill_manual(values = c("coral2", "dodgerblue", "goldenrod2", "lightblue", "springgreen", 
            "grey90", "grey80", "grey70", "grey60"))

ggsave("recommendations_bar.pdf", height = 12, width = 20, dpi = 300)
```
